---
title: "week7"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Executables: 
* Rewrite algorithm to rank users against each other
    + use business interests research from week6
    + try different ranking variable combinations to see how they compare
    
    ie:
    masked_id  | sav_bal | checking_cnt | ... | loyalty_rank 
    ------------- | ------------- | ------------- | ------------- | -------------
    1  | 10 | 23 | ...| 16.5
    2  | ... | ... | ... | ...
    ... | ... | ... | ... | ...
    
* Scatterplot ranked users against other fields (ie.age) to look for corrlations 
## Variables to rank
* checking_acct_ct: WF makes money from account fees
* savings_acct_ct: WF makes money from account fees
* sav_bal_altered: WF makes money off of loans they give using the money customers keep in their savings accounts 
* atm_withdrawals_cnt: WF makes money from atm fees and overdraft fees
* online_bank_cnt: Research shows that customers are loyal to banks with usable web pages
* direct_phone_cnt: Research shows that customers are loyal to banks that offer human interaction services
## Variables to compare
* age
* tenure_altered
* mortgage_flag
* heloc_flag
* personal_loan_flag
* cc_flag
* branch_visit_cnt
* phone_banker_cnt
## Import Data
```{r}
library(readxl)
month_end_balances <- read_excel("/usr/local/Learn2Mine-Main/galaxy-dist/lesson_datasets/Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
    sheet = "Month end balances ", col_types = c("numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))
```
## Select month of December 2016
```{r}
# create data frame to manipulate
df <- data.frame(month_end_balances)
# select only rows where asof_yyyymm = 201612
df_12 <- df[df$asof_yyyymm==201612,]
```
## Rank Users by each variable
```{r}
# Rank by checking_acct_ct, savings_acct_ct, sav_bal_altered, atm_withdrawls_cnt, online_bank_cnt, direct_phone_cnt
df_rank <- df_12
df_rank$checking_acct_ct_rank <- rank(df_rank$checking_acct_ct, ties.method= "average")
df_rank$savings_acct_ct_rank <- rank(df_rank$savings_acct_ct, ties.method= "average")
df_rank$sav_bal_altered_rank <- rank(df_rank$sav_bal_altered, ties.method= "average")
df_rank$atm_withdrawls_cnt_rank <- rank(df_rank$atm_withdrawls_cnt, ties.method= "average")
df_rank$online_bank_cnt_rank <- rank(df_rank$online_bank_cnt, ties.method= "average")
df_rank$direct_phone_cnt_rank <- rank(df_rank$direct_phone_cnt, ties.method= "average")
# Take row average and store in Loyalty_num
df_rank$loyalty_num <- rowMeans(df_rank[,c("checking_acct_ct_rank","savings_acct_ct_rank","sav_bal_altered_rank","atm_withdrawls_cnt_rank","online_bank_cnt_rank","direct_phone_cnt_rank")])
# Rank Loyalty_num and store in Loyalty_rank
df_rank$loyalty_rank <- rank(df_rank$loyalty_num, ties.method="average")
```
## Scatterplots
These scatterplots are intended to show correlations between a customer's rank and other variables, if a correlation exists. None can be detected from the loyalty rank algorithm version 1.
```{r}
#
library(ggplot2)
ggplot(df_rank, aes(x=loyalty_rank, y=age)) + geom_point()
ggplot(df_rank, aes(x=loyalty_rank, y=tenure_altered)) + geom_point()
ggplot(df_rank, aes(x=loyalty_rank, y=mortgage_flag)) + geom_point()
ggplot(df_rank, aes(x=loyalty_rank, y=heloc_flag)) + geom_point()
ggplot(df_rank, aes(x=loyalty_rank, y=personal_loan_flag)) + geom_point()
ggplot(df_rank, aes(x=loyalty_rank, y=cc_flag)) + geom_point()
ggplot(df_rank, aes(x=loyalty_rank, y=branch_visit_cnt)) + geom_point()
ggplot(df_rank, aes(x=loyalty_rank, y=phone_banker_cnt)) + geom_point()
```
## Prediction
Error rate is way too high: 98%.
```{r}
library(randomForest)
fit <- randomForest(as.factor(loyalty_rank) ~ age + tenure_altered + mortgage_flag + heloc_flag + personal_loan_flag + cc_flag + branch_visit_cnt + phone_banker_cnt,
                      data=df_rank, 
                      importance=TRUE, 
                      ntree=2000)
varImpPlot(fit)



