---
title: "week8"
author: "Paul Anderson"
date: "10/9/2017"
output: html_document
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Signing up for an account
Because we are using a server behind a Firewall, you'll need to publish your apps in order to see them. Not a problem. It's all free and easy to use. 

[Instructions to get things published](https://shiny.rstudio.com/articles/shinyapps.html)

Once you create your account, you'll want to run code that looks like the following but with your own values of course:

rsconnect::setAccountInfo(name='pauleanderson',
 		  token='31AD5F9D543A5763DF9291D1FA4F991C',
		  secret='<SECRET>')

## Generating a graph for a user
## Getting the data ready

This is all copied from the previous markdown.
```{r context="data"}
library(readxl)
library(randomForest)

Fake_Data_and_Metadata_Final_no_pass <- read_excel("Fake+Data+and+Metadata+-+Final+no+pass.xlsx", 
                                                   sheet = "Daily use of a WF credit card ")

library(reshape)
reshaped_df = cast(Fake_Data_and_Metadata_Final_no_pass, masked_id ~ Des1, value = 'Payment', fun.aggregate=sum)
# when we do this it is a good idea to fix the names of the columns
tidy.name.vector <- make.names(colnames(reshaped_df), unique=TRUE)
colnames(reshaped_df) = tidy.name.vector

reshaped_df[is.na(reshaped_df)] = 0
rownames(reshaped_df) = reshaped_df$masked_id
reshaped_df = reshaped_df[,-1] # Remove the first column

reshaped_df_norm = t(apply(reshaped_df, 1, function(x) x/sum(x))) # divide every row by the sum of that row
# this returns a matrix without column names, so we need to add them back and change back to a data frame
colnames(reshaped_df_norm) = colnames(reshaped_df)
reshaped_df_norm = as.data.frame(reshaped_df_norm)

# Pre-compute everything because we want to go fast
masked_ids = c(1,2,3)

results = list()
row_ids = list()
for (j in 1:length(masked_ids)) {
  masked_id = masked_ids[j]
  row_id = which(rownames(reshaped_df_norm)==masked_id)
  train.data = reshaped_df_norm[-row_id,]
  # Remove that user from the models
  masked_id_fits = list() # Store the results in a list for examination later
  for (i in 1:ncol(reshaped_df_norm)) {
    masked_id_fits[[i]] <- randomForest(train.data[,i] ~ .,
                        data=train.data,
                        importance=TRUE, 
                        ntree=2000)
  }
  # This is where we save things for later
  results[[j]] = masked_id_fits
  row_ids[[j]] = row_id
}
```


## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r context="render", echo=FALSE}
library(plotly)

inputPanel(
  selectInput("masked_id", label = "Masked ID:",
              choices = masked_ids, selected = masked_ids[1])
)
plotOutput("distPlot")

plotlyOutput("barPlot")

```

```{r context="server", echo=FALSE}
library(randomForest)

output$distPlot <- renderPlot({
  masked_id = input$masked_id
  j = which(masked_ids == masked_id)
  row_id = row_ids[[j]]
  masked_id_fits = results[[j]]
  
  predictions = c()
  for (i in 1:ncol(reshaped_df_norm)) {
    predictions[i] = predict(masked_id_fits[[i]],reshaped_df_norm[row_id,])
  }
  errors = abs(predictions - as.vector(as.matrix(reshaped_df_norm[row_id,])))
  
  inxs = order(errors)
  
  hist(as.vector(errors[inxs]))
})

library(plotly)
output$barPlot <- renderPlotly({
  masked_id = input$masked_id
  j = which(masked_ids == masked_id)
  row_id = row_ids[[j]]
  masked_id_fits = results[[j]]
  
  predictions = c()
  for (i in 1:ncol(reshaped_df_norm)) {
    predictions[i] = predict(masked_id_fits[[i]],reshaped_df_norm[row_id,])
  }
  errors = abs(predictions - as.vector(as.matrix(reshaped_df_norm[row_id,])))
  
  inxs = order(errors)
  
  plot_ly(
    x = 100*as.vector(errors[inxs]),
    y = factor(colnames(reshaped_df_norm)[inxs],levels = colnames(reshaped_df_norm)[inxs]),
    name = "Errors",
    type = "bar"
  )
})
```

## Running this example
You can execute the following:

rmarkdown::run("week8.Rmd", shiny_args=list(host="0.0.0.0",port=6000))

BUT you must change the port number to the one I gave your group, and then go to:

https://freyja.cs.cofc.edu/6000